---
title: "World Politics"
subtitle: "Alliances and Security Cooperation"
author: "Michael E. Flynn"
institute: "Kansas State University"
date: "`r Sys.Date()`"
output:
  xaringan::moon_reader:
    css: ["default", "lecture_theme.css"]
    nature:
      highlightStyle: github
      highlightLines: true
      ratio: 16:9
      countIncrementalSlides: false
      titleSlideClass: [middle, left, inverse]
---

```{r setup, include = FALSE, echo=FALSE, warning=FALSE}
library(ggplot2)
library(data.table)
library(peacesciencer)
library(tidyverse)
library(janitor)
library(here)
library(ghibli)
library(ggforce)

knitr::opts_chunk$set(comment = '', dpi = 400, fig.width = 8, fig.height = 5, warning = FALSE, echo = FALSE, message = FALSE)

sysfonts::font_add_google("Oswald", family = "oswald")
showtext::showtext_auto()

# Windows and Mac seem to process the fotn differently
if (Sys.info()[[1]] == "Windows"){

basesize <- 35

} else {
  
basesize <-  11
  
}

update_geom_defaults("text", list(size = 10, family = "oswald", fontface = "plain", lineheight = 0.2))


# Regular plot theme
theme_flynn <- function(){ 
  
      theme_linedraw(base_size = basesize, base_family = "oswald") %+replace% 
        
        theme(plot.title = element_text(face = "bold", size = basesize * 1.3, hjust = 0, margin = margin(t = 0, b = 0.2, l = 0, r = 0, unit = "cm")),
              plot.subtitle = element_text(size = basesize, hjust = 0, margin = margin(t = 0.0, b = 0.2, unit = "cm")),
              plot.caption = element_text(face = "italic", size = basesize * 0.6, hjust = 1, lineheight = 1, margin = margin(t = 0.2, unit = "cm")),
              panel.border = element_rect(fill = NA, size = 0.2),
              strip.background = element_rect(fill = "gray80", color = "black", size = 0.2),
              strip.text = element_text(size = basesize, color = "black", face = "bold", margin = margin(t = 0.2, b = 0.2, l = 0.2, r = 0.2, unit = "cm")),
              panel.background = element_rect(size = 0.2),
              panel.grid.major = element_line(color = "gray70", size = 0.15),
              panel.grid.minor = element_line(color = "gray90", size = 0.1),
              axis.title = element_text(face = "bold", size = basesize),
              axis.title.y = element_text(angle = 90, margin = margin(t = 0, r = 0.2, b = 0, l = 0, unit = "cm")),
              axis.title.x = element_text(margin = margin(t = 0.2, r = 0, b = 0, l = 0, unit = "cm")),
              axis.ticks = element_line(size = 0.1),
              axis.ticks.length = unit(0.1, "cm"),
              legend.title = element_text(size = basesize, face = "bold", hjust = 0, margin = margin(t = 0, b = 0, l = 0, r = 0, unit = "cm")),
              plot.margin = margin(0.5, 0.5, 0.5, 0.5, unit = "cm"),
              legend.margin = margin(t=0, b=0, r=0, l=0),
              legend.box.margin = margin())
}

#install.packages("RefManageR")

#library(RefManageR)
#library(bibtex)
#BibOptions(check.entries = FALSE,
#           bib.style = "authoryear", # Bibliography style
#           max.names = 3, # Max author names displayed in bibliography
#           sorting = "nyt", #Name, year, title sorting
#           cite.style = "authoryear", # citation style
#           style = "markdown",
#           hyperlink = FALSE,
#           dashed = FALSE)

#myBib <- ReadBib(here::here("../../classrefs.bib", check = FALSE))

```


# Lecture Overview

1. Alliances—what are they?

2. Patterns in alliances over time

3. Why and when do states form alliances?

4. Other forms of security cooperation



---

# Key Questions and Goals

1. Develop an understanding of the factors that drive states to cooperate militarily in general

2. Develop an understanding of the factors that drive more powerful states to ally with weaker states, specifically

3. Define and understand the security-autonomy trade off

4. Develop an understanding of how other issue areas, like trade, investment, and institutions, relate to alliance relationships


---
class: inverse, center, middle

# Defining Alliances



---
# Defining Alliances

We talk about alliances a lot, but what are they?

> **Alliance** Institutions that help their members cooperate militarilty in the event of a war (FLS)

- Note that Alliances are *formal* agreements

- States with common interests and policy goals can coordinate action without formally defining their obligations

- This means they're *aligned* but not necessarily *allied*





---
# Defining Alliances

Frieden, Lake, and Schultz offer a basic definition, but some additional detail might be helpful:

Alliances Treaty Obligations and Provisions (ATOP) definition:

> Written agreements, signed by official representatives of at least two independent states, that includ promises to aid a partner in the event of military conflict, to remain neutral in the event of conflict, to refrain from military conflict with one another, or to consult/cooperate in the even tof international crises that create the potential for military conflict (Leeds, Ritter, Mitchell, and Long 2002: 238).



---
# Defining Alliances

ATOP's definition includes five alliance pact types:

1. Defense

2. Offense

3. Neutrality

4. Nonaggression

5. Consultation




---
class: center, middle

```{r alliance-all}

# need to build basic timeline data frame because not all types are present in all years
time <- data.frame(highally = c("Defense", "Offense", "Neutrality", "Non-Aggression", "Consultation")) %>% 
  mutate(year = rep(list(seq(1815, 2018)), 5)) %>% 
  unnest(year)

# Note that alliances still in effect have endyr coded 0. Don't know why. This seems like a silly coding method.
allies <- fread(here("../../../Data Files/ATOP 5_0 (csv)/atop5_0a.csv")) %>% 
  group_by(atopid) %>% 
  mutate(endyr = ifelse(endyr == 0, 2018, endyr),
         year = list(seq(begyr, endyr))) %>% 
  dplyr::select(atopid, year, defense, offense, neutral, nonagg, consul, bilat) %>% 
  mutate(provisions = sum(defense, offense, neutral, nonagg, consul),
         highally = case_when(
           defense == 1 ~ "Defense",
           defense == 0 & offense == 1 ~ "Offense",
           defense == 0 & offense == 0 & neutral == 1 ~ "Neutrality",
           defense == 0 & offense == 0 & neutral == 0 & nonagg == 1 ~ "Non-Aggression",
           defense == 0 & offense == 0 & neutral == 0 & nonagg == 0 & consul == 1 ~ "Consultation",
         )) %>% 
    unnest(year) 


ally.year <- time %>% 
  left_join(allies) %>% 
  group_by(year, highally) %>% 
  dplyr::summarise(allycount = n_distinct(atopid)) %>% 
  mutate(highally = factor(highally, levels = c("Defense", "Offense", "Neutrality", "Non-Aggression", "Consultation")))




#ghibli::ghibli_palettes$MarnieDark1

ggplot(ally.year, aes(x = year, y = allycount, group = highally)) +
  geom_area(fill = "#2F1619FF") +
  theme_flynn() +
  scale_x_continuous(breaks = seq(1815, 2020, 15)) +
  labs(x = "Year",
       y = "Count",
       title = "Alliances in the international system, 1815-2016",
       subtitle = "Where do we see major changes in the number of alliances?",
       caption = "Data obtained from the Alliance Treaty Obligations and Provisions (ATOP) Data: http://www.atopdata.org/")

  
```



---
class: center, middle

```{r alliance-disaggregated}

ggplot(ally.year, aes(x = year, y = allycount, group = highally, fill = highally)) +
  geom_area(color = "white", size = 0.2) +
  theme_flynn() +
  scale_x_continuous(breaks = seq(1815, 2020, 15)) +
  viridis::scale_fill_viridis(option = "magma", discrete = TRUE) +
  labs(x = "Year",
       y = "Count",
       fill = "Commitment Type",
       title = "Alliances in the international system, 1815-2016",
       subtitle = "Where do we see major changes in the number of alliances?",
       caption = "Data obtained from the Alliance Treaty Obligations and Provisions (ATOP) Data: http://www.atopdata.org/")

```




---
class: center, middle

```{r alliance-disaggregated-percent}

ggplot(ally.year, aes(x = year, y = allycount, group = highally, fill = highally)) +
  geom_area(position = "fill", color = "white", size = 0.2) +
  theme_flynn() +
  scale_x_continuous(breaks = seq(1815, 2020, 15), expand = c(0,0)) +
  scale_y_continuous(labels = scales::percent, expand = c(0,0)) +
  viridis::scale_fill_viridis(option = "magma", discrete = TRUE) +
  labs(x = "Year",
       y = "Count",
       fill = "Commitment Type",
       title = "Alliances in the international system, 1815-2016",
       subtitle = "Where do we see major changes in the number of alliances?",
       caption = "Data obtained from the Alliance Treaty Obligations and Provisions (ATOP) Data: http://www.atopdata.org/")

```