---
title: "World Politics"
subtitle: "Challenges to the Post-War Order"
author: "Michael E. Flynn"
institute: "Kansas State University"
date: "`r Sys.Date()`"
header-includes:
  - \usepackage{tikz}
  - \usepackage{pgfplots}
output:
  xaringan::moon_reader:
    css: ["default", "lecture_theme.css"]
    nature:
      highlightStyle: github
      highlightLines: true
      ratio: 16:9
      countIncrementalSlides: false
      titleSlideClass: [middle, left, inverse]
---
  
```{r setup, include = FALSE, echo=FALSE, warning=FALSE}
#remotes::install_github("gadenbuie/xaringanExtra")

library(ggplot2)
library(data.table)
library(peacesciencer)
library(kableExtra)
library(tidyverse)
library(janitor)
library(here)
library(kableExtra)
library(ggpattern)
library(ghibli)
library(ggforce)
library(patchwork)
library(scales)
library(rmapshaper)
library(sf)
library(sp)
library(tmap)
library(raster)
library(broom)
library(xaringanExtra)
library(gghighlight)
library(tidyquant)

xaringanExtra::use_scribble()
xaringanExtra::use_editable(expires = 365)

knitr::opts_chunk$set(comment = '', dpi = 400, fig.width = 8, fig.height = 5, warning = FALSE, echo = FALSE, message = FALSE)

sysfonts::font_add_google("Oswald", family = "oswald")
showtext::showtext_auto()

# Windows and Mac seem to process the fotn differently
if (Sys.info()[[1]] == "Windows"){
  
  basesize <- 35
  
} else {
  
  basesize <-  11
  
}

update_geom_defaults("text", list(size = 10, family = "oswald", fontface = "plain", lineheight = 0.2))


# Regular plot theme
theme_flynn <- function(){ 
  
  theme_linedraw(base_size = basesize, base_family = "oswald") %+replace% 
    
    theme(plot.title = element_text(face = "bold", size = basesize * 1.3, hjust = 0, margin = margin(t = 0, b = 0.2, l = 0, r = 0, unit = "cm")),
          plot.subtitle = element_text(size = basesize, hjust = 0, margin = margin(t = 0.0, b = 0.2, unit = "cm")),
          plot.caption = element_text(face = "italic", size = basesize * 0.6, hjust = 1, lineheight = 0.3, margin = margin(t = 0.2, unit = "cm")),
          panel.border = element_rect(fill = NA, size = 0.2),
          strip.background = element_rect(fill = "gray80", color = "black", size = 0.2),
          strip.text = element_text(size = basesize, color = "black", face = "bold", margin = margin(t = 0.2, b = 0.2, l = 0.2, r = 0.2, unit = "cm")),
          panel.background = element_rect(size = 0.2),
          panel.grid.major = element_line(color = "gray70", size = 0.15),
          panel.grid.minor = element_line(color = "gray90", size = 0.1),
          axis.title = element_text(face = "bold", size = basesize),
          axis.title.y = element_text(angle = 90, margin = margin(t = 0, r = 0.2, b = 0, l = 0, unit = "cm")),
          axis.title.x = element_text(margin = margin(t = 0.2, r = 0, b = 0, l = 0, unit = "cm")),
          axis.ticks = element_line(size = 0.1),
          axis.ticks.length = unit(0.1, "cm"),
          legend.title = element_text(size = basesize, face = "bold", hjust = 0, margin = margin(t = 0, b = -0.3, l = 0, r = 0, unit = "cm")),
          legend.text = element_text(size = basesize, hjust = 0, margin = margin(t = 0, b = 0, l = -0.5, r = 0, unit = "cm")),
          legend.box.spacing =  unit(0.2, "cm"),
          plot.margin = margin(0.3, 0.3, 0.3, 0.3, unit = "cm"))
}

#install.packages("RefManageR")

#library(RefManageR)
#library(bibtex)
#BibOptions(check.entries = FALSE,
#           bib.style = "authoryear", # Bibliography style
#           max.names = 3, # Max author names displayed in bibliography
#           sorting = "nyt", #Name, year, title sorting
#           cite.style = "authoryear", # citation style
#           style = "markdown",
#           hyperlink = FALSE,
#           dashed = FALSE)

#myBib <- ReadBib(here::here("../../classrefs.bib", check = FALSE))

```


# Lecture Overview

1. Background Review

2. Emerging Challenges

3. Moving Forward



---
class: inverse, center, middle

# Background Review




---
# Background Review

Four primary "pillars" of the international order

1. Sovereignty

2. Collective Security

3. Bretton Woods System

4. Promotion of Human Rights and Democracy

--

These things are based on and implemented using

- Liberal values

- Multilateral cooperation

- US Hegemony


---
# Background Review

Key Institutions

- United Nations

- General Agreement on Tariffs and Trade (World Trade Organization)

- International Monetary Fund

- World Bank

- North Atlantic Treaty Organization (NATO)




---
class: center, middle

```{r, dpi=400}

bases <- troopdata::basedata

map <- rnaturalearth::ne_countries(returnclass = "sf") %>% filter(name != "Antarctica")

ggplot() +
  geom_sf(data = map, aes(group = name, geometry = geometry)) +
  geom_point(data = bases, aes(x = lon, y = lat), color = "darkblue", alpha = 0.7) +
  theme_void() +
  theme(text = element_text(family = "oswald", face = "bold")) +
  labs(title = "US military facilities around the world, 1950-2015")


```




---
class: center, middle, inverse

# Emerging Challenges



---
class: left, top

# Emerging Challenges


What are the issues?

- (Bad) Non-state actors

- Nuclear proliferation

- Biological weapons

- Cyberattacks

- Revisionist powers

- Global health and pandemics

- Climate change and environmental degradation 

- Populist movements 

- Political and economic instability



---
class: center, middle

```{r}

nuclearpowers <- c("United States", "United Kingdom", "Russia", "China", "France", "Pakistan", "India", "North Korea", "Israel")

map <- rnaturalearth::ne_countries(returnclass = "sf") %>% 
  mutate(nuclear = ifelse(name %in% nuclearpowers, "Yes", "No")) %>% 
  filter(name != "Antarctica")

ggplot() +
  geom_sf(data = map, aes(geometry = geometry, fill = nuclear), color = "black", size = .1) +
  viridis::scale_fill_viridis(discrete = TRUE, begin = 0.25, end = 0.85) +
  theme_void() +
  theme(text = element_text(family = "oswald", face = "bold")) +
  labs(title = "States possessing nuclear weapons",
       fill = "Possesses\nNuclear Weapons")


```




---
class: center, middle

```{r}

gdp <- peacesciencer::create_stateyears(system = "gw") %>% 
  add_sdp_gdp() %>% 
  filter(year >= 1950) %>% 
  dplyr::rename("ccode" = "gwcode")

capabilities <- peacesciencer::cow_nmc %>% 
  filter(year >= 1950) %>% 
  left_join(gdp, by = c("ccode", "year"))


ggplot(capabilities %>% filter(year == 2015), aes(x = exp(wbgdp2011est), y = cinc)) +
  geom_point(alpha = 0.75, shape = 21, fill = "orange", size = 3) +
  ggrepel::geom_label_repel(aes(label = statename)) +
  theme_flynn() +
  scale_x_continuous(labels = scales::dollar_format()) +
  labs(x = "World Bank GDP Estimate",
       y = "CINC Score",
       title = "GDP and Correlates of War CINC Scores, 2015",
       caption = "Data obtained from the {peacesciencer} package.")

```



---
class: center, middle

```{r}


ggplot(capabilities %>% filter(year == 2015), aes(x = exp(wbgdp2011est), y = milex)) +
  geom_point(alpha = 0.75, shape = 21, fill = "orange", size = 3) +
  ggrepel::geom_label_repel(aes(label = statename)) +
  theme_flynn() +
  scale_x_continuous(labels = scales::dollar_format()) +
  scale_y_continuous(labels = scales::dollar_format()) +
  labs(x = "World Bank GDP Estimate",
       y = "Military Expenditures (thousands of USD)",
       title = "GDP and military expenditures, 2015",
       caption = "Data obtained from the {peacesciencer} package.")
```



---
class: center, middle, inverse

# Moving Forward



---
class: left, top

# Moving Forward

So what's to be done?

- Lots of different issues, but there are also lots of common themes and characteristics

- The work is increasingly interconnected, and so externalities from human behavior (both positive and negative) are increasingly visible

- People have shared interests in solving these problems

- Useful frameworks to keep in mind:

  - Bargaining (think about people's interests)
  
  - Cooperation and coordination problems (think about rules and incentive structures)
